The Ecology of Cloud Computing

### 2022-04-12 paper.tex
Fixed typos, fixed citations and added diagrams

\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage[numbers,sort]{natbib}
\usepackage{hyperref}

@@

@@

An NFT uses an algorithm to create non-fungible digital data - like a digital certificate of authenticity. It uses a technology called blockchain to make a digital receipt that cannot be copied or faked. A blockchain is essentially a list (of transactions) and a signature - the signature is a line of letters and numbers that encapsulates the contents of the list. If someone tampers with the list, or tries to improperly add something to it, the signature and list will not match and you will be able to tell it has been tampered with.\citep{narayanan2017bitcoin} This can be a good thing; artists have found use in it, being able to sell digital art as a limited-stock item (like traditional paintings are) directly, without than having to go through a middleman like an auction house. In fact, in 2020,  artists earned \$1.8 million \citep{redman2020NFT} from selling their art as NFTs.\newline

@@

\subsection{Data}
Once we have processed our data, what do we do with it? Data storage counts for around 10\% of data center energy usage, for a total of 20TWh per year \citep{shehabi2016united} - nearly 40x the energy used by all personal computers in the UK in a year.  \citep{waters2019energy} Not to mention the energy used to produce all this storage media - the manufacture of a Dell PowerStore9000T hard drive (one of their largest available) produces approximately 2765kg CO2e. \citep{dell2021powerstore} A possible and quite successful solution used for data archival is to move unused data to magnetic tapes; with careful discernment of what data needs to be at one's fingertips versus on a library shelf. Using tape can reduce the space and utilities needed for storage sixfold. \citep{moore2007disk}. \newline

@@

While machine learning has been an amazing development in several fields, it runs the risk of overuse; some software concepts can almost become fashionable, with not enough consideration put into them. Note that the statistics in the previous paragraph were for \emph{training} a model - that is, we generated over 5kg CO2e and \emph{not even solved any problems yet}. Compared to a non-machine learning approach, it is entirely possible that this lump sum of carbon right at the start will put its footprint far ahead of a non-learning-based software for essentially its entire lifetime - care must be taken that machine learning is the right tool for the job.

@@

The PUE of a data centre can be reduced through various considerations; the first is how to manage heat emissions. Cooling the servers using water cooling rather than air conditioning (in fact, air conditioning has recently come under fire for creating a vicious cycle as the energy from air conditioning heats the planet, which means more air conditioning is required, so on \citep{ospina2018cooling}) can be much more efficient. Design can be used to further improve this; a common way of cooling data centres is by cooling alternating rows of servers while letting the odd rows heat up, which creates a convection current around the whole room that keeps everything cool. The water cooling itself can also be made more effective using ground (passive) cooling - by piping the water underground before piping it back into the data centre, we can take advantage of the effectively unlimited cold under the earth. \newline

We can also integrate some interesting computing approaches to infrastructure to reduce PUE. \citet{evans2016deepmind} reduced the energy use for cooling at a Google datacentre by 40\% through training an AI on centre infrastructure data such as temperatures, power, pump speeds (for water cooling) and then allowing it to control the datacentre's cooling system. However, for 'smaller wins', we must consider the power use of our computing approaches; the actual energy use of training and running this AI may be non-negligible (although in this case, probably less than 40\% of the cooling bill) \newline

@@

\subsection{Moving costs}
When we move our data over to a cloud centre, we must choose how we move it, as well as how we move large data back and forth from user to server in future. Thus, time for an interesting comparsion; the carbon use of data transfer versus that of a heavy-goods vehicle. Huge corporations like Google transfer data using 'SneakerNet' - that is, loading hard drives into a van and taking them physically to their destination.\footnote{A wonderful article comparing SneakerNet to traditional bandwidth here: \href{https://what-if.xkcd.com/31/}{https://what-if.xkcd.com/31/}} As long as travel time is under 57 hours, SneakerNet is more efficient than the internet in terms of transfer speed \citep{jaya2020sneakernet} - for reference, crossing the United States takes 40-50 hours. But what about carbon? \newline

>>>>

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage[numbers,sort]{natbib}
\usepackage{hyperref}

@@

\begin{figure}[!ht]
\begin{center}
    \includegraphics[width=.6\textwidth]{Diagrams/GLOBE---1850-2021-MO-withlabels.png}
        \caption{A graphic of global average temperature per year from 1850 to 2021 (I believe this image should be somewhere in every paper on anthropogenic climate change). \citep{hawkins2021show}}
\end{center}
\end{figure}

@@

An NFT uses an algorithm to create non-fungible digital data - like a digital certificate of authenticity. It uses a technology called blockchain to make a digital receipt that cannot be copied or faked. A blockchain is essentially a list (of transactions) and a signature - the signature is a line of letters and numbers that encapsulates the contents of the list. If someone tampers with the list, or tries to improperly add something to it, the signature and list will not match and you will be able to tell it has been tampered with.\citep{narayanan2017bitcoin} This can be a good thing; artists have found use in it, being able to sell digital art as a limited-stock item (like traditional paintings are) directly, without having to go through a middleman (e.g. an auction house). In fact, in 2020,  artists earned \$1.8 million \citep{redman2020NFT} from selling their art as NFTs.\newline

@@

\subsection{Data}
Once we have processed our data, what do we do with it? Data storage counts for around 10\% of data center energy usage, for a total of 20TWh per year \citep{shehabi2016united} - nearly 40x the energy used by all personal computers in the UK in a year.  \citep{waters2019energy} Moreover, large amounts of energy is needed to produce all this storage media - the manufacture of a Dell PowerStore9000T hard drive (one of their largest available) produces approximately 2765kg CO2e. \citep{dell2021powerstore} A possible and quite successful solution used for data archival is to move unused data to magnetic tapes; with careful discernment of what data needs to be at one's fingertips versus on a library shelf. Using tape can reduce the space and utilities needed for storage sixfold. \citep{moore2007disk}. \newline

@@

While machine learning has been an amazing development in several fields, it runs the risk of overuse; some software concepts can almost become fashionable, with not enough consideration put into them. Note that the statistics in the previous paragraph were for \emph{training} a model - that is, we generated over 5kg CO2e and \emph{have not even solved any problems yet}. Compared to a non-machine learning approach, it is entirely possible that this lump sum of carbon right at the start will put its footprint far ahead of a non-learning-based software for essentially its entire lifetime - care must be taken that machine learning is the right tool for the job.

@@

The PUE of a data centre can be reduced through various considerations; the first is how to manage heat emissions. Cooling the servers using water cooling rather than air conditioning (in fact, air conditioning has recently come under fire for creating a vicious cycle as the energy from air conditioning heats the planet, which means more air conditioning is required, and so on \citep{ospina2018cooling}) can be much more efficient. Design can be used to further improve this; a common way of cooling data centres is by cooling alternating rows of servers while letting the odd rows heat up, which creates a convection current around the whole room that keeps everything cool. The water cooling itself can also be made more effective using ground (passive) cooling - by piping the water underground before piping it back into the data centre, we can take advantage of the effectively unlimited cold under the earth. \newline

We can also integrate some interesting computing approaches to infrastructure to reduce PUE. \citet{evans2016deepmind} reduced the energy use for cooling at a Google datacentre by 40\% through training an AI on centre infrastructure data such as temperatures, power, pump speeds (for water cooling) and then allowing it to control the datacentre's cooling system. However, for 'smaller wins', we must consider the power use of our computing approaches; the actual energy use of training and running this AI may be non-negligible (although in this case, probably less than 40\% of the cooling bill). \newline

@@

\subsection{Moving costs}
When we move our data over to a cloud centre, we must choose how we move it, as well as how we move large data back and forth from user to server in future. Thus, time for an interesting comparison; the carbon use of data transfer versus that of a heavy-goods vehicle. Huge corporations like Google transfer data using 'SneakerNet' - that is, loading hard drives into a van and taking them physically to their destination.\footnote{A wonderful article comparing SneakerNet to traditional bandwidth here: \href{https://what-if.xkcd.com/31/}{https://what-if.xkcd.com/31/}} As long as travel time is under 57 hours, SneakerNet is more efficient than the internet in terms of transfer speed \citep{jaya2020sneakernet} - for reference, crossing the United States takes 40-50 hours. But what about carbon? \newline
